---
title: "Flight Data Analysis"
---

## Overview

This page contains the detailed analysis of flight data from the Department of Transportation.

## Data Loading

```{r}
#| echo: true
#| eval: false
#| label: load-data

# Load necessary libraries
library(readr)
library(dplyr)
library(ggplot2)

# Load your CSV files
# Replace 'example_flight_data.csv' with your actual CSV file names
flight_data <- read_csv("data/example_flight_data.csv")

# Display first few rows
head(flight_data)

# Basic summary
summary(flight_data)
```

**Note**: The code blocks above are set to `eval: false` by default. Change to `eval: true` when you're ready to execute them with your actual data.

## Key Findings

### Finding 1: Flight Delays by Carrier

Analyze which carriers have the most delays.

```{r}
#| echo: true
#| eval: false
#| label: finding-1-delays

# Calculate average delay by carrier
delay_summary <- flight_data %>%
  group_by(carrier) %>%
  summarise(
    avg_delay = mean(delay_minutes, na.rm = TRUE),
    total_flights = n()
  ) %>%
  arrange(desc(avg_delay))

# Visualize
ggplot(delay_summary, aes(x = reorder(carrier, avg_delay), y = avg_delay)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Average Delay by Carrier",
    x = "Carrier",
    y = "Average Delay (minutes)"
  ) +
  theme_minimal()
```

### Finding 2: Flight Routes Analysis

Examine the most common routes and their performance.

```{r}
#| echo: true
#| eval: false
#| label: finding-2-routes

# Create route identifier and analyze
route_analysis <- flight_data %>%
  mutate(route = paste(origin, "→", destination)) %>%
  group_by(route) %>%
  summarise(
    flight_count = n(),
    avg_delay = mean(delay_minutes, na.rm = TRUE)
  ) %>%
  arrange(desc(flight_count))

# Display top routes
head(route_analysis, 10)

# Visualize top routes
route_analysis %>%
  head(10) %>%
  ggplot(aes(x = reorder(route, flight_count), y = flight_count)) +
  geom_col(fill = "coral") +
  coord_flip() +
  labs(
    title = "Top 10 Flight Routes",
    x = "Route",
    y = "Number of Flights"
  ) +
  theme_minimal()
```

### Finding 3: Flight Status Distribution

Analyze the distribution of flight statuses (on-time vs delayed vs cancelled).

```{r}
#| echo: true
#| eval: false
#| label: finding-3-status

# Count flights by status
status_summary <- flight_data %>%
  group_by(status) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

# Create a pie chart or bar chart
ggplot(status_summary, aes(x = status, y = count, fill = status)) +
  geom_col() +
  geom_text(aes(label = paste0(count, "\n(", percentage, "%)")), 
            vjust = -0.5) +
  labs(
    title = "Flight Status Distribution",
    x = "Status",
    y = "Number of Flights"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Summary Statistics

```{r}
#| echo: true
#| eval: false
#| label: summary-stats

# Overall summary statistics
cat("Total number of flights:", nrow(flight_data), "\n")
cat("Average delay time:", mean(flight_data$delay_minutes, na.rm = TRUE), "minutes\n")
cat("Median delay time:", median(flight_data$delay_minutes, na.rm = TRUE), "minutes\n")

# Detailed summary
summary(flight_data)

# Additional statistics by carrier
flight_data %>%
  group_by(carrier) %>%
  summarise(
    flights = n(),
    avg_delay = mean(delay_minutes, na.rm = TRUE),
    max_delay = max(delay_minutes, na.rm = TRUE),
    min_delay = min(delay_minutes, na.rm = TRUE)
  )
```

## Conclusions

[Add your conclusions based on the analysis]

## Next Steps

[Optional: Add any future work or additional analysis planned]
